### 2019-01-29 ambiguity/simulation-runners/discourse-ambiguity.Rmd

---
title: "Discourse-ambiguity"
author: "Ben"
date: "11/9/2018"
output: html_document
---

Simulation 2 run code for "The interactions of rational, pragmatic agentslead to efficient language structure and use." Note that we make use of the `rwebppl` package to interface with webppl. Modeling and discourse generation are contained in the "ambiguity/webppl" folder.

```{r libraries, message=FALSE, warning=FALSE}
rm(list = ls())  # clear workspace

library(cowplot)
library(doParallel)
library(entropy)
library(ggthemes)
library(tidyverse)
library(zoo)

source(file.path(getwd(), "../utils.R"))
```

Load model.
```{r getModel}
fPath <- file.path(getwd(), '../webppl/discourse_ambiguity_model.wppl')
m <- getModelFile(fPath)
```

Setup.
```{r runFn}
# Single run function
runAmbiguity <- createRunFn(m)
runAmbiguityFn <- function(i, targetDistr, nUtterances, resultType, alpha, recursionLevel) {
  dTemp <- data.frame(targetDistr=targetDistr, 
                      nUtterances=nUtterances, 
                      resultType=resultType, 
                      alpha=alpha, 
                      recursionLevel=recursionLevel)
  df <- runAmbiguity(dTemp) %>%
    mutate(runNum=i, targetDistr=targetDistr, alpha=alpha, resultType=resultType, recursionLevel=recursionLevel)
  df
}

# Multiple runs function
runAmibguityCorpus <- function(topic, alpha, resultType, recursionLevel, nUtterances, nSims) {
  # Parallelization setup
  no_cores <- detectCores() - 1
  cl <- makeCluster(no_cores, type='FORK')
  registerDoParallel(cl)

  # Run sims
  ptm <- proc.time()
  sims <- foreach(i=seq(1, nSims), .packages=c('dplyr', 'rwebppl'), .combine=rbind) %dopar% 
    runAmbiguityFn(i,
                   targetDistr=topic,
                   nUtterances=nUtterances,
                   resultType=resultType,
                   alpha=alpha,
                   recursionLevel=recursionLevel)
  stopCluster(cl)
  etm <- proc.time() - ptm

  # Preprocess for entropy calculations.
  binSize <- 1
  df_binned <- addUtteranceBins(sims, nUtterances=nUtterances, binSize=binSize)
  df_filled <- fillUtteranceProportions(df_binned, binSize) %>%
    mutate(recursionLevel=recursionLevel)
  list(sims, df_filled)
}
```

# Run sims
```{r}
# Note to use cached results you must hace access to the data. Data is available on request bpeloqui@stanford.edu.
fpPreprocssed <- "/Users/benpeloquin/Data/zipf_principles/discourse_ambiguity/df_preprocessed_20190124.csv"
fpSims <- "/Users/benpeloquin/Data/zipf_principles/discourse_ambiguity/df_simsCmbd_20190124.csv"

topic <- 'T1'
nUtterances <- 30
alpha=8
nSims <- 100

simsBaseline <- runAmibguityCorpus(
  topic=topic,
  alpha=8,
  resultType='baseline',
  recursionLevel=1,
  nUtterances=nUtterances,
  nSims=nSims)

simsUnaware <- runAmibguityCorpus(
  topic=topic,
  alpha=8,
  resultType='discourseUnaware',
  recursionLevel=1,
  nUtterances=nUtterances,
  nSims=nSims)

simsAware <- runAmibguityCorpus(
  topic=topic,
  alpha=8,
  resultType='discourseAware',
  recursionLevel=1,
  nUtterances=nUtterances,
  nSims=nSims)

simsAmbiguityCorpusUnaware <- simsUnaware[[1]]
dfAmbiguityCorpusUnaware <- simsUnaware[[2]]
simsAmbiguityCorpusAware <- simsAware[[1]]
dfAmbiguityCorpusAware <- simsAware[[2]]
simsAmbiguityCorpusBaseline <- simsBaseline[[1]]
dfAmbiguityCorpusBaseline <- simsBaseline[[2]]

# Combine data
simsCmbd <- rbind(simsAmbiguityCorpusUnaware,
                  simsAmbiguityCorpusAware,
                  simsAmbiguityCorpusBaseline)
# Write raw sims data
write.csv(simsCmbd, fpSims)

dfAmbiguityCorpusCmbd <- rbind(dfAmbiguityCorpusUnaware,
                               dfAmbiguityCorpusAware,
                               dfAmbiguityCorpusBaseline)

# Pre-process
preProcessedSims <- simsCmbd %>% 
  group_by(resultType, alpha) %>%
  mutate(isAmbiguous=(utterance=='x' | utterance=='y'),
         isX=utterance=='x',
         isA=utterance=='a',
         isB=utterance=='b') %>%
  group_by(resultType, alpha, utteranceNum) %>%
  summarise(n=n(),
            numAmb=sum(isAmbiguous),
            probAmb=numAmb/n,
            stdAmb=sqrt(probAmb*(1-probAmb)/n),
            numX=sum(isX),
            probX=numX/n,
            stdX=sqrt(probX*(1-probX)/n),
            numA=sum(isA),
            probA=numA/n,
            stdA=sqrt(probA*(1-probA)/n),
            numB=sum(isB),
            probB=numB/n,
            stdB=sqrt(probB*(1-probB)/n),
            avgT1=mean(T1),
            sdT1=sd(T1),
            avgT2=mean(T2),
            sdT2=sd(T2),
            avgT3=mean(T3),
            sdT3=sd(T3),
            avgT4=mean(T4),
            sdT4=sd(T4))
# Cache data
write.csv(preProcessedSims, fpPreprocssed)
df_preProcessedSims <- preProcessedSims
```