### 2019-01-31 paper/paper.Rmd
update objective function -- integrate comments from NDG about emphasis on novelty of objective and our framing.

---
title: "The interactions of rational, pragmatic agents  \n lead to efficient language structure and use"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Benjamin N. Peloquin} \\ \texttt{bpeloqui@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Noah D. Goodman} \\ \texttt{ngoodman@stanford.edu} \\ Department of Computer Science \\ Stanford University
    \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@university.edu} \\ Department of Psychology \\ Stanford University}

abstract: 
    Languages display a diverse set of distributional regularities such as the relation between a word's frequency and rank in a corpus, the distribution of syntactic dependency lengths across languages, or the presence of properties such as ambiguity. We discuss a framework for studying how these system-level properties emerge from local, in-the-moment interactions of rational, pragmatic speakers and listeners. To do so, we derive a novel objective function for measuring the communicative efficiency of linguistic systems in terms of the interactions of speakers and listeners. We examine the behavior of this objective in a series of simulations focusing on the communicative function of ambiguity in language. These simulations suggest that rational pragmatic agents will produce communicatively efficient systems and that interactions between such agents provide a framework for examining efficient properties of language structure and use more broadly.

keywords:
    "Communicative efficiency, Rational Speech Act theory, computational modeling, information theory, agent-based simulation"

output: cogsci2016::cogsci_paper
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(gridExtra)
library(ggplot2)
library(xtable)
```

# Introduction

Why do languages look the way they do? Zipf [-@Zipf1949a] proposed that distributional properties found in natural language were evidence of speaker-listener effort minimization. In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence for this claim has been largely derived at the level of the lexicon. Zipf argued that the particular relationship between a word’s frequency and its rank, length, and denotation size could be explained as an emergent property of speaker-listener effort minimization. \par

Zipf articulated what is now considered a *functionalist* approach to language science -- analyzing language structure and use in terms of efficiency. Such an approach might reframe our opening question as follows: how does having property \textit{x} make using language $\ell$ more or less useful for communication? Among others, this efficiency-based framing has produced a rich set of theoretical and empirical targets exploring semantic typology [@RegierKempKay2015a], properties such as ambiguity [@Piantadosi2011a] and compositionality [@KirbyGriffithsSmith2014a], and the efficient use of reduction and redundancy in production [@LevyJaeger2007a; @GenzelCharniak2002a]. These projects typically employ information-theoretic principles to characterize efficiency in natural or simulated language corpora. We adopt a different approach in this work, deriving a measure of efficiency from first principles of \textit{efficient language use} and then showing how this can lead to communicatively efficient systems.\par

Functionalist theories commonly frame language efficiency in terms of a fundamental effort-assymetry underlying everyday communication: what is “hard” for a speaker is likely different than what is “hard” for a listener. Zipf gave the following intution for how this idea -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of a single, low-cost word. Given such a language, the full set of potential meanings would be conveyed using only that word, i.e. $\ell_{speaker}^*$ would be fully ambiguous and all possible meanings would need to be disambiguated by a listener. From the standpoint of listener effort, an optimal language $\ell_{listener}^*$ would map all possible meanings to distinct words, removing a listener's need to disambiguate. In this example, speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*. Clearly, natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$. Zipf proposed that the particular lexicon-level properties he observed were a result of these competing forces -- the pressure to jointly minimize speaker and listener effort.\par

But how does this optimization take place? The example given by Zipf [-@Zipf1949a], describes local, communicative interactions in terms of a \textit{reference game}. Speakers intend to refer to some object in the world $m$. They choose some utterance $u$ to transmit this intended meaning, $u \rightarrow m$. The listener attempts to reconstruct this intended meaning given the transmitted utterance, $m \rightarrow u$. Other projects have assumed this basic reference game setting [@RegierKempKay2015a; @Piantadosi2011a] and this simplification of the communicative act has proven productive in theoretical [@FerreriCancho2018a], simulation-based [@KirbyGriffithsSmith2014a] and empirical explorations [@HawkinsFrankeSmithGoodman2018a] of efficient language structure and use.\par

Zipf's conception of speaker and listener effort may be connected to speakers' pragmatic reasoning in the moment. Under a Gricean treatment of pragmatics, speakers and listeners follow a set of conversational maxims in which they cooperate to transfer information [@Grice1975a]. These maxims appear to emerge from efficiency concerns, however (Horn, 1984). We formalize this connection -- showing how system-level efficiencies can emerge from local interaction behavior. Intuitively, our claim is that to understand an "efficient" property of a system it is essential that we consider how that property is *used* efficiently. It follows that any quantitative measure of "efficiency" must be a function of both the system and its users.\par

In the current work we provide a case-study for the framing presented above. We choose a property of languages that could, in principle, vary freely, but shows strong regularities across languages. The explanadum is why this regularity holds. We examine ambiguity as our property, extending ideas by Piantadosi et al. [-@Piantadosi2011a]. We define a novel measure of efficiency that depends on the interactional behavior of speaker and listener agents. We adopt the reference game as our primary unit of interaction and model language users using the Rational Speech Act (RSA) framework -- a computional model of language use, which is nicely attested by experimental data on interaction. Using these basic ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency. Further, we show how ambiguity is \textit{used} efficiently during local, in-the-moment interactions. The contributions of this work are twofold -- we derive a novel objective for measuring linguistic efficiency and also show how the reference game framework, in combination with formal models of communication, can be used to connect ideas about system-level efficiencies to in-the-moment language use.\par

# Exploring efficient language design and use in rational pragmatic agents

```{r plot-reference-game, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2.5, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "An example reference game with associated literal semantics (in our terminology a ``language'')."}
img <- png::readPNG("figs/game1.png")
grid::grid.raster(img)
```

### Reference games
Zipf’s example of optimal speaker- and listener-languages took the form of a reference game. We adopt that formulation here, assuming these communication games as our basic unit of analysis. In this framework, speakers and listeners are aware of a set of objects $M$ (*meanings*) and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1). Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$. Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}. We consider a set of contexts $c\in C$ with an associated prior $P(C)$. Each context $c_i$ describes a different distribution over meanings e.g. $p(M|C=c_i) \neq p(M|C=c_j)$. Finally, we consider a set of communicative events $e \in E$ where $<u, m, c> = e$ is an utterance, meaning and context triple.\par

### Languages
A language $\ell$ defines the set of semantic mappings between utterance and meanings. For example, Figure 1 contains four utterances $U = \{\text{blue}, \text{green}, \text{square}, \text{circle}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}$. The boolean matrix describes the literal semantics of the language. We define a language as "ambiguous" if there is some utterance $u \in U$ which can apply to multiple meanings (i.e. $|[[u_i]]| > 1$)\footnote{We use double brackets $[[\dots]]$ to represent denotation.}. In Figure 1 both the words "square" and "green" are ambigous so we would say that $\ell$ contains ambiguity.

### Speakers and listeners
The Rational Speech Act framework (RSA) is computational-level theory of pragmatic language use, which has been well attested in experimental work examining both speaker and listener communication behavior [@FrankGoodman2012a; @GoodmanFrank2016a]. RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a]. We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting\footnote{See SI section 1 for a more detailed description of basic RSA speaker-listener defintions and to Goodman \& Frank [-@GoodmanFrank2016a] for an overview of its applications.}.\par

An RSA *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using $\ell$ in a given context $c$. That is, a speaker defines $P_{speaker}(u|m, c; \ell)$. We will use $S(u|m, c; \ell)$ as short-hand throughout.  A *listener agent* defines a conditional distribution over meanings, mapping from utterances $U$ to meanings $M$ using $\ell$ in a given context $c$ (i.e. $L(m|u, c;\ell)$). Note that both speakers and listeners can induce joint distributions over utterace-meaning pairs, although, importantly, these distributions may differ:
$$P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)$$
$$P_{listener}(u, m| c; \ell) = L(m|u, c; \ell)p(u|c)$$

# Zipfian objective for linguistic system efficiency

Zipf (1949) proposed that the particular distributional properties found in natural language emerge from competing speaker and listener pressures. We operationalize this objective in equation (1) -- the efficiency of a linguistic system $\ell$ being used by speaker and listener agents $S$ and $L$ is the sum of the expected speaker and listener effort to communicate over all possible communicative events $e \in E$.\par

\begin{equation}
\begin{split}
  \text{Efficiency}(S, L, \ell) = \mathbb{E}_{e \sim P(E)}[\text{speaker effort}] \\+ \mathbb{E}_{e \sim P(E)}[\text{listener effort}]
\end{split}
\end{equation}

We assume that speaker effort is related to the suprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterances costs as independent from context (ie. $p(u|c)p(c)=p(u)p(c)$).} -- intuitively, the number of bits needed to encode the utterance $u$. This particular formalization of speaker-cost is general enough to accommodate a range of cost instantiations, such as production difficult via articulation effort, cognitive effort related to lexical access, or others [@BennettGoodman2015a].\par

$$\text{speaker effort} = -log_2(p(u|c))$$

We assume listener effort is the semantic suprisal of a meaning given an utterance. This operationalization of listener effort is intuitively related to existing work in sentence processing in which word comprehension difficulty is proportional to surprisal [@Hale2001a; @Levy2008a].

$$\text{listener effort} = -log_2(L(m|u, c; \ell))$$

Importantly, we assume that events $e = <u, m, c>$ are sampled according the to following generative model -- some context occurs in the world with probability $P(C=c)$. Within this context, an object $m$ occurs with probability $p(m|c)$. The speaker attempts to refer to that object by sampling from her conditional distribution $S(u|m, c; \ell)$ (i.e. $e \sim p(c)p(m|c)S(u|m, c; \ell)$). From the ingredients described above, it is possible to derive the following objective between the speaker and listener joint distributions (see SI 2.2 for a complete derivation).

\begin{equation}
\begin{split}
  = \mathbb{E}_{c \sim P(C)}[H_{cross}(P_{speaker}, P_{listener} | c; \ell)]\\
\end{split}
\end{equation}
From an information-theoretic perspective this objective is intuitive: The Cross-Entropy (CE) gives us a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs. In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$. In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listeners think.

# Simulating the communicative function of ambiguity

The task of understanding language is marked by a frequent need to handle various forms of ambiguity: lexical, syntactic, among others[@WasowPerforsBeaver2005a]. The ubiquity of this property, however, has been argued to provide evidence that langauges are not been optimized for communication\footnote{Chomsky's (2002) claim that "... language is not well designed, because you have such properties as ambiguity" is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$.} [@Chomsky2002a].

Piantadosi et al. [-@Piantadosi2011a] argue just the opposite, claiming that ambiguity is an *efficient* property of any communication system in which *communication is contextualized*. Simply put, it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low. In particular, context (or common ground) can provide useful information for disambiguation.\par

We re-iterate their argument here. Say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), which differ in cost, and two languages ($\ell_1$ and $\ell_2$) which describe different utterance-meaning mappings. In language $\ell_1$, the low-cost $u_1$ can be used to refer to both $m_1$ and $m_2$ ($[[u_1]]_{\ell_1} = \{m_1, m_2 \}$), but the high-cost $u_2$ cannot be used at all ($[[u_2]]_{\ell_1} = \emptyset$). By contrast, in language $\ell_2$, $u_1$ can only refer to $m_1$ and $u_2$ can only refer to $m_2$ ($[[u_1]]_{\ell_2} = \{m_1\}$ and $[[u_2]]_{\ell_2} = \{m_2 \}$). While it is cheaper for a speaker to use $\ell_1$ (because speaking is always low-cost), it is more difficult for a listener (because $u_1$ is ambiguous). Crucially, however, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred to $\ell_2$.\par

In the following experiments we explore two aspects of Piantadosi’s claim. First, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ is most likely to contain ambiguous lexical items. In the second experiment, we explore an efficient language *use* aspect of the claim -- at what point in a conversation is it useful for a speaker to use ambiguous lexical material?\par

```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Optimal languages are more likely to contain ambiguous items as the amount of contextual information increases. Vertical axis shows the proportion of optimal languages containing ambiguity. Horizontal axis shows the number of context-sizes (1-4). Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only objectives. Error bars represent 95 percent confidence intervals."}
img <- png::readPNG("figs/fig1.png")
grid::grid.raster(img)
```

# Simulation 1:  Optimal languages contain ambiguity when context is informative

We adopt the hypothesis presented by Piantadosi, formalizing their argument in our current framework -- showing that ambiguity is an efficient property under our CE objective in the reference game setting. We proceed by generating languages with different amounts of contextual support (varying the size of $|C|$). We search the space of languages, examining whether ones which minimize our objective contain ambiguity. If context leads to more efficient commnication, then optimal languages under our objective should be more likley to be ambiguous as the amount of context increases.\par

## Simulation set-up

We conduct $N=2000$ simulations. For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$ (recall that $U$ is our set of utterances and $M$ our set of meanings). Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every column has some non-zero assignment). For a given simulation the goal is to find the language $\ell^*$ which minimizes our CE objective and then check to see if that language contains ambiguity.\par

We define language efficiency as a function of the particular semantic mappings induced by that language, the speaker and listener agents ($S$ and $L$), as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$). Rather than assume particular structure, for each simulation we generate $P(U) \sim \text{Dir}(1, |U|)$, $P(M|C=c) \sim \text{Dir}(1, |M|)$ (a separate conditional distribution over meanings for each context $c$), and $P(C) \sim \text{Dir}(1, |C|)$, where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.\par

### Context

Following the argument given by Piantadosi et al. [-@Piantadosi2011a], we want to assess the impact of *context* on the presence of ambiguity in optimal languages. To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M|C), P(C)\}$. Our first is a *one-context* condition ($|C|=1$) -- only a single distribution over meanings $P(M)$. In our two-context condition ($|C| = 2$), we consider efficiency under both $P(M|C=c_1)$ as well as $P(M|C=c_2)$. Three- and four-context conditions corresponding accordingly.\par

### Baselines

For comparison, we also examine properties of optimal languages under two additional objectives. Zipf [-@Zipf1949a] proposed that the speaker-optimal language $\ell_{speaker}^*$ would minimize speaker effort and the listener-optimal language $\ell_{listener}^*$ would minimize listener effort. We define these objectives using the first and second half of equation 1 (see SI section 2.2. for more detail).\par

## Results and Discussion

Figure 1. plots the proportion of optimal languages under each objective as a function of context size. The red line shows that as the number of contexts increases, so does the probability that an optimal language $\ell^*_{cross}$ contains ambiguity under our CE objective. For comparison we also plot the proportion of speaker-optimal $\ell^*_{speaker}$ (blue line) and listener-optimal $\ell^*_{listener}$ (green line) languages that contain ambiguity. In line with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages always contain ambiguity. If languages are designed to minimize listener effort then ambiguity is always avoided.\par

In simulation 1 we explored the degree to which ambiguity was an efficient property when communication was contextualized. We showed that as contextual information increased, the optimal languages under our CE objective were more likely to have ambiguous lexical items. Importantly, we did not see this effect under alternative objectives which only emphasized speaker- or listener effort alone.\par

While our results indicate that ambiguity is an efficient property of contextualized language, these simulations assumed that agents could always disambiguate items in context. That is, in our four conditions both agents had perfect knowledge of the relevant conditional distributions ($P(M|C)$). This assumption may be too strong for describing much of day-to-day communication -- we seldom interact with others with perfect knowledge of the current context (or topic) of conversation. To explore how ambiguity may be \textit{used} efficiencly, we next explore a case in which the listener has imperfect knowledge of context at the start of the conversation, but may infer it from a discourse history in Simulation 2.\par

# Simulation 2: Rational, pragmatic speakers use ambiguity efficiently

We shift focus in Simulation 2 -- instead of examining the landscape of a space of languages under our CE objective, we consider a single language containing ambiguity, examining how that property is used over a discourse. If context is not informative with respect to meaning early in a discourse, but is at later positions, then an efficient speaking strategy would avoid ambiguous material early, saving it for when contextual information is disambiguating. In this toy example we can consider “context” as analogous to a “topic” of conversation. As an example, imagine a scenario in which a reader is beginning a magazine article. While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved. In this setting, using an potential low-cost, but ambiguous pronoun early may lead to misunderstanding if context is sufficiently informative. But, if by a later position enough contextual information has accumulated, it may be efficient for the writer to use the ambiguous material.\par

##  Simulation set-up

We consider consider a single language $\ell$, which contains both ambiguous and unambiguous utterances. We assume ambiguous utterances are lower cost. Crucially, we do not assume that the listener knows the particular topic ($c_{current}$) of the conversation \textit{a priori}. Rather, that the listener has knowledge of the set of possible topics $C = \{c_1, \dots, c_k\}$, but \textit{does not know which one is currently being used by the speaker}. Formally, this means the listener does not have access to the correct conditional distribution over meanings $P(M|C=c_{current})$ at the start of the discourse.\par

Over the course of a discourse $D$, the listener tries to infer both the current topic, $c_{current}$, as well as the particular meaning $m$ of a given utterance $u$. That is, we consider updated listener and speakers ($L(m, c|u,D;\ell)$, $S(u|m,c,D;\ell)$) in which both agents can track the history of previous utterances $D$. Importantly, an agent can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$ (see SI section 3 for more detail).\par

```{r plot-optimal-use, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "(A) shows the empirical probability that our speaker used an ambiguous utterance as a function of discourse position. (B) shows speaker effort across the three models. (C) shows the Cross-Entropy objective under our three speaker models. Error bars represent 95 percent confidence intervals."}

img <- png::readPNG("figs/discourse_grid_plot.png")
grid::grid.raster(img)
```

We conduct $N=600$ simulations, generating discourses of length $|D|=30$ utterances, comparing three speaker models ($n=200$ each). We consider a single language $\ell$\footnote{See SI for the matrix notation of this langauge.} with $|U|=6$ and $|M|=4$ in which two of the utterances are ambiguous and lower cost than the unambiguous utterances. (Note that use of this particular language is not essential -- the results are broadly generalizable to languages that contain ambiguity.)\par

## Speaker agents
We consider three types of speaker models. Our \textit{Full pragmatics} agent, models a speaker who reasons about her listener and also has complete recall of the set of utterances in the discourse $D$. This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discrouse, but should infer it over the discourse. We compare two baselines models. The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to discourse history. The second, a \textit{No pragmatics} speaker who does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see SI section 3 for details).\par

## Hypotheses

We are interested in \textit{when} it is efficient to use low-cost, ambiguous material in this discourse setting. A speaker strategy that is mutually efficient for both agents should avoid ambiguity until sufficient contextual information has accumulated. We should expect this to be reflected in our \textit{Full pragmatics} model who reasons about the listener and discourse history. By contrast, a speaker-optimal model who does not consider the listener should greedily use ambiguous utterances (\textit{No Pragmatics} model), while a listener-optimal model should avoid ambiguity entirely (\textit{Partial pragmatics} model).\par

## Results and Discussion

Figure 2, (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position. The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse and the \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely. But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds. (C) tracks our CE objective for each model over the discourse. Note that the objective decreases for all three models, primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$). However, the objective declines more quickly under the \textit{Full} and \textit{Partial} pragmatic speakers as listener agents are better able to infer the correct context. Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs described in (B). While speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines in the \text{Full Pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par

In Simulation 2 we explored how ambiguity could be employed efficiently during language use. We assumed that the listener did not know the current context (topic) \textit{a priori}, but could infer it from the discourse history. We examined speaker production behavior, tracking use of low-cost, but ambiguous material over the discourse. Early in the discourse the \textit{Full pragmatic} speaker avoided using ambiguous material, only using the material later when the current topic was known to the listener.

# General Discussion

How do the competing pressures of speakers and listeners give rise to the distributional forms found in natural language? Zipf [-@Zipf1949a] proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon. We explored the interactions of rational pragmatic agents as a framework for understanding efficient language structure and use. We focused on an argument on the communicative function of ambiguity [@Piantadosi2011a], deriving a novel speaker-listener Cross-Entropy objective for measuring the efficiency linguistic systems from first principles of efficient language use. In a series of simulations we showed that optimal languages are more likely to contain ambiguous material when context is informative and that rational pragmatic agents will *use* ambiguous material efficiently.\par

A limitation of the current work is an analysis of exactly how this objective compares to existing measures. For example, previous work has described competing speaker-listener pressures in terms of a trade-off of \textit{simplicity} and \textit{informativeness} [@KempRegier2012a] or \textit{expressivity} and \textit{compressibility} [@SmithTamarizKirby2013a]. Future work should assess the degree to which we can derive the same properties as previous studies using our current framework. In line with this idea, a primary focus of the current work is to provide a case-study in the productivity of assessing functionalist theories in repeated reference games between rational pragmatic agents. In theory, this framework can serve as a domain general tool to assess the range of functionalist theories examining efficient language-structure and use.\par

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
 \noindent