### 2019-02-01 ambiguity/simulation-runners/context-ambiguity-plots.Rmd
initial ndg changes to simulation 2 intro and matrix added to plot 1. just pushing in case i mess all the formatting up...

---
title: "ambiguity_context_analysis"
author: "Ben"
date: "11/12/2018"
output: html_document
---

```{r libraries}
rm(list = ls())  # clear workspace
library(cowplot)
library(tidyverse)
library(ggthemes)
```

```{r read-files}
fp <- '/Users/benpeloquin/Data/zipf_principles/ambiguity/runs-20190121/uniform_uniform_1_1_500.csv'
fp <- '/Users/benpeloquin/Desktop/Projects/zipf_principles/test-20190130/uniform_uniform_1_1_500.csv'
fp <- '/Users/benpeloquin/Data/zipf_principles/ambiguity/runs-20190130/uniform_uniform_1_1_500.csv'
df <- read.csv(fp)
```

## Claim 1 - the more contexts the more useful ambiguity is.
```{r plot1-a}
p1 <- df %>%
  select(sim_id, idxs, n_contexts, speaker_k,
         utterance_order_fn_name, is_min_ce,
         is_min_listener_entropy, is_min_speaker_entropy, 
         contains_ambiguity) %>%
  gather(objective, val, c(is_min_ce, is_min_listener_entropy, is_min_speaker_entropy)) %>%
  # filter(objective=='base_speaker_listener_ce') %>%
  mutate(val=as.logical(val),
         contains_ambiguity=as.logical(contains_ambiguity)) %>%
  filter(val) %>%  # only look at 'optimal languages'
  group_by(objective, n_contexts) %>%
  summarise(n=n(),  # number of optimal languages (there can be ties)
            prop_ambiguous=mean(contains_ambiguity),
            # Binomial CIs
            ci_high = prop_ambiguous + qnorm(0.975)*sqrt((prop_ambiguous*(1 - prop_ambiguous))/n),
            ci_low=prop_ambiguous - qnorm(0.975)*(sqrt((prop_ambiguous*(1-prop_ambiguous))/n))) %>%
  ungroup() %>%
  # Normal approximation isn't great when we have proportions so 
  # close to 1. so just rounding down slightly. This is a little
  # loosey-goosey, but does not impact the overall narrative
  # in a significant way
  # mutate(ci_high=ifelse(ci_high > 1, 1, ci_high)) %>%
  mutate(objective=ifelse(objective=='is_min_ce', 'Cross-entropy',
                          ifelse(objective=='is_min_listener_entropy', 'Optimal Listener', 'Optimal Speaker'))) %>%
  rename(Objective=objective) %>%
  ggplot(aes(x=n_contexts, y=prop_ambiguous, col=Objective)) +
    geom_line(size=1, alpha=0.9) +
    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), width=0.05, alpha=0.7, col='black') +
    ylab("Percent optimal languages \n containing ambiguity") +
    xlab("Number of contexts (|C|)") +
    theme_few() +
    # In case we want to embed legend in the plot, but not really necessary...
    theme(legend.position = c(0.7, 0.3),
          legend.background = element_rect(colour = 'grey', fill = 'white', linetype='solid'),
          plot.margin = unit(c(0, 0, 0, 0), "cm"))
```

```{r plot1-b}
p2 <- df %>%
  select(sim_id, idxs, n_contexts, is_min_ce, ce, base_speaker_listener_ce, contains_ambiguity) %>%
  filter(as.logical(is_min_ce)) %>%
  mutate(ce_minus_base=ce-base_speaker_listener_ce) %>%
  ungroup %>%
  group_by(n_contexts) %>%
  summarise(n=n(),
            diff=mean(ce_minus_base),
            ci_high=diff + 1.96*sd(ce_minus_base)/sqrt(n),
            ci_low=diff - 1.96*sd(ce_minus_base)/sqrt(n)) %>%
  ggplot(aes(x=n_contexts, y=diff)) +
    geom_bar(stat='identity', position='dodge', alpha=0.8) +
    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), width=0.1, col='black', alpha=0.8) +
    theme_few() +
    ylab("Difference in cross-entropy\n aware vs unaware of context") +
    xlab("Number of contexts (|C|)")
```

Saves plot grid with example unambiguous and ambiguous language.
```{r}
p2 <- ggdraw() +
  draw_image("../../paper/figs/ambiguous_lang.png", scale=0.75)
p3 <- ggdraw() +
  draw_image("../../paper/figs/unambiguous_lang.png", scale=0.75, clip="on")

g1 <- plot_grid(p2, p3, align='vh', ncol=1, labels=c("B", "C")) +
  theme(plot.margin = unit(c(0, 0, 0, 0.4), "cm"))
g2 <- plot_grid(p1, g1, ncol=2, labels=c("A", "", ""))

ggsave(file.path(getwd(), "../../paper/figs/fig1.png"), plot=g2, width=7, height=3)
```

Saves single fig plot.
```{r saving}
# plot_grid(p1, p2, labels = c("A", "B"), align = "h")
# ggsave(file.path(getwd(), "../../paper/figs/fig1.png"), width=10, height=3)
ggsave(file.path(getwd(), "../../paper/figs/fig1.png"), plot=p1, width=7, height=3)
```

Save grid plot
```{r saving}
# plot_grid(p1, p2, labels = c("A", "B"), align = "h")
# ggsave(file.path(getwd(), "../../paper/figs/fig1.png"), width=10, height=3)
```
```{r}
df %>%
  filter(sim_id==1, n_contexts==1) %>%
  select(idxs, is_min_ce, is_min_listener_entropy, is_min_speaker_entropy) %>%
  gather(typ, val, c(is_min_ce, is_min_listener_entropy, is_min_speaker_entropy)) %>%
  filter(val=='True')

```